### 2020-03-28 thesis/paper.tex
variable names

\subsection{DBSCAN}
Density-Based Spatial Clustering of Application with Noise (DBSCAN) \cite{10.5555/3001460.3001507} is an algorithm, that unlike k-Means, is not based on partitioning the data directly but on searching for dense regions in the data. For the calculation, two parameters are needed, namely epsilon ($\epsilon$) and a minimum number of objects ($\mu$ or MinPts). The parameter $\epsilon$ defines how large a region may be considered for neighborhood search and $\mu$ defines how many objects must be within that region for it to be considered dense. The following explanations are needed for a full understanding of the algorithm:
\newline

@@

\subsection{OPTICS}\label{sec:opticstheory}
Ordering Points To Identify the Clustering Structure or OPTICS \cite{10.1145/304181.304187} is an extension of DBSCAN whereby DBSCAN's Epsilon $\epsilon$ parameter is used in a different manner. It also has a parameter MinPts which has a similar meaning as the $\mu$ parameter of DBSCAN. The idea is that different distance parameters are used in each step such that density-based clusters with different densities can be found simultaneously. To achieve this, the  $\epsilon$ parameter only defines the maximum distance for which points can be considered reachable and the actual reachability distance for two objects is computed as:
$$reachability\text{-}distance_{\epsilon,MinPts}(p,o)=$$
$$
\begin{cases}
    \text{UNDEFINED},& \text{if } |N_\epsilon(p)|< \text{MinPts}\\
    max(core\text{-}distance(o),distance(o,p)),              & \text{otherwise}
\end{cases}
$$
whereby:
$$core\text{-}distance_{\epsilon,MinPts}(p)=$$
$$
\begin{cases}
    \text{UNDEFINED},& \text{if } |N_\epsilon(p)|< \text{MinPts}\\
    MinPts\text{-}distance(p),              & \text{otherwise}
\end{cases}
$$

@@

Sampling 200 runs of DBSCAN by fixing minPTS to $5$ and randomly picking $\epsilon$ between $0.01$ and $0.5$ (which is roughly the range where non-trivial clusterings can result) gives us a best result with an NMI of $0.896$. It can be seen in Figure \ref{fig:zahn_best}. The points colored in white in Figure \ref{fig:zahn_best} indicate points that DBSCAN considered noise, for the calculation of the NMI and in all other parts of the tool they are considered to belong to a common cluster.

@@

To generate base clusterings k-Means was used, varying the value for $k$ between $2$ and $10$ and computing $10$ samples for each value. Trivial solutions were omitted and no ground truth was added as none is available. With the workflow defined, it is executed also performing OPTICS for meta-clustering using the default MinPTS and $\epsilon$. With the ordering resulting from the OPTICS algorithm, the heat map for the resulting clusterings is created and can be seen in Figure \ref{fig:multi_heatmap}.

>>>>

\subsection{DBSCAN}
Density-Based Spatial Clustering of Application with Noise (DBSCAN) \cite{10.5555/3001460.3001507} is an algorithm, that unlike k-Means, is not based on partitioning the data directly but on searching for dense regions in the data. For the calculation, two parameters are needed, namely epsilon ($\epsilon$) and a minimum number of objects ($\mu$ or $minPts$). The parameter $\epsilon$ defines how large a region may be considered for neighborhood search and $\mu$ defines how many objects must be within that region for it to be considered dense. The following explanations are needed for a full understanding of the algorithm:
\newline

@@

\subsection{OPTICS}\label{sec:opticstheory}
Ordering Points To Identify the Clustering Structure or OPTICS \cite{10.1145/304181.304187} is an extension of DBSCAN whereby DBSCAN's Epsilon $\epsilon$ parameter is used in a different manner. It also has a parameter $minPts$ which has a similar meaning as the $\mu$ parameter of DBSCAN. The idea is that different distance parameters are used in each step such that density-based clusters with different densities can be found simultaneously. To achieve this, the  $\epsilon$ parameter only defines the maximum distance for which points can be considered reachable and the actual reachability distance for two objects is computed as:
$$reachability\text{-}distance_{\epsilon,minPts}(p,o)=$$
$$
\begin{cases}
    \text{UNDEFINED},& \text{if } |N_\epsilon(p)|< minPts\\
    max(core\text{-}distance(o),distance(o,p)),              & \text{otherwise}
\end{cases}
$$
whereby:
$$core\text{-}distance_{\epsilon,minPts}(p)=$$
$$
\begin{cases}
    \text{UNDEFINED},& \text{if } |N_\epsilon(p)|< minPts\\
    minPts\text{-}distance(p),              & \text{otherwise}
\end{cases}
$$

@@

Sampling 200 runs of DBSCAN by fixing $minPTS$ to $5$ and randomly picking $\epsilon$ between $0.01$ and $0.5$ (which is roughly the range where non-trivial clusterings can result) gives us a best result with an NMI of $0.896$. It can be seen in Figure \ref{fig:zahn_best}. The points colored in white in Figure \ref{fig:zahn_best} indicate points that DBSCAN considered noise, for the calculation of the NMI and in all other parts of the tool they are considered to belong to a common cluster.

@@

To generate base clusterings k-Means was used, varying the value for $k$ between $2$ and $10$ and computing $10$ samples for each value. Trivial solutions were omitted and no ground truth was added as none is available. With the workflow defined, it is executed also performing OPTICS for meta-clustering using the default $minPTS$ and $\epsilon$. With the ordering resulting from the OPTICS algorithm, the heat map for the resulting clusterings is created and can be seen in Figure \ref{fig:multi_heatmap}.
